---
title: "PA1 Assignment"
author: "Mohamed Soliman"
date: "August 9, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Code for reading in the dataset, Installing dependencies and processing the data  

```{r, results='asis'}
dataFile <- unzip(zipfile = "activity.zip")
activity <- read.csv(dataFile, stringsAsFactors=FALSE)
activity$date <- as.Date(activity$date)
if(!require(dplyr)) {install.packages("dplyr") ; library(dplyr)}
if(!require(ggplot2)) {install.packages("ggplot2") ; library(ggplot2)}
```


## 2. Histogram of the total number of steps taken each day

```{r}
total_steps_per_day <- activity %>% group_by(date) %>% summarise(total_steps = sum(steps))

ggplot(data=total_steps_per_day, aes(x=date, y=total_steps)) + geom_bar(stat="identity")
```


## 3. Mean and median number of steps taken each day

```{r}
mean_steps <- mean(total_steps_per_day$total_steps, na.rm = TRUE)

median_steps <- median(total_steps_per_day$total_steps, na.rm = TRUE)
```

The Mean Steps is `r mean_steps`, and the Median is `r median_steps`.


## 4. Time series plot of the average number of steps taken

```{r}
avg_steps_per_interval <- activity %>% group_by(interval) %>% summarise(mean_steps = mean(steps, na.rm = TRUE))

ggplot(data=avg_steps_per_interval, aes(x=interval,y=mean_steps)) + geom_line()
```


## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max_interval <- as.integer(avg_steps_per_interval[which(avg_steps_per_interval == max(avg_steps_per_interval)), 1])
```


## 6. Code to describe and show a strategy for imputing missing data

```{r}
total_missing <- sum(activity$steps)
```

Number of rows with missing value `r total_missing`. 

We are going to replace the missing steps by the mean for that 5-minute interval.

```{r}
means <- avg_steps_per_interval$mean_steps
imputed_activity <- activity

for (inter in avg_steps_per_interval$interval){
 index <- which(avg_steps_per_interval$interval== inter)
 imputed_activity <- imputed_activity %>% mutate(steps=replace(steps, interval==inter & is.na(steps),means[index])) 
}
```


## 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
total_steps_per_day_imputed <- imputed_activity %>% group_by(date) %>% summarise(total_steps = sum(steps))

ggplot(data=total_steps_per_day_imputed, aes(x=date, y=total_steps)) + geom_bar(stat="identity")

imputed_mean_steps <- mean(total_steps_per_day_imputed$total_steps)

imputed_median_steps <- median(total_steps_per_day_imputed$total_steps)
```

The Mean Steps is `r imputed_mean_steps`, and the Median is `r imputed_median_steps`.


## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r pressure, echo=FALSE}

imputed_activity <- imputed_activity %>% mutate(weekdays = ifelse(weekdays(date) %in% c("Saturday","Sunday") ,"weekend" , "weekday" ))

imputed_activity$weekdays <- as.factor(imputed_activity$weekdays)

avg_steps_per_interval_imputed <- imputed_activity %>% group_by(weekdays, interval) %>% summarise(mean_steps = mean(steps))

ggplot(data=avg_steps_per_interval_imputed, aes(x=interval,y=mean_steps, group=1)) + geom_line() + facet_grid(weekdays ~ .)
```


